\chapter*{Curriculum	Vitae}

\textbf{PROFILE}

Software engineer with over 18+ years experience in Software Engineering, Business to Business applications, Object Oriented design and methodologies. Highly Proficient in C/C++, UNIX and .NET platform. Up to date knowledge of distributed systems, Algorithms, Computational Mathematics, transaction processing, relational database and network engineering. Strong distributed system design, system programming background, excellent knowledge of mission critical applications, networks, protocols. 

\textbf{KEY SKILLS}

    • Operating Systems: UNIX (Linux, UNIX , Windows .NET,  Win32, Embedded Systems

    • Methodologies and concepts: Object Oriented analysis and design, Client/Server and N-tier programming, Design of Distributed  Systems

    • Real life experience of scaling and optimizing large systems, designing for reliability and scalability

    • Throughout knowledge of C++, STL programming practices and paradigms. Early adopter of new C++ standards such as C++03, C++11.

    • Excellent knowledge of full range of Microsoft technologies, and recent expertise in  C\# ,.NET,  ASP.NET platform and SQL server (complete .NET stack), Excel RTD Server and Excel Extensions

    • Deep knowledge and experience with: distributed systems, Web services, concurrent programming, threads, real time programming,  embedded systems, TCP/IP and network protocols

    • Experience with database theory, database programming, ORACLE 9i and 10g administration, mySQL

    • Implemented Classification, Data analytics,  Machine Learning solutions, using NumPY, Tensor Flow and Keras

    • Theoretical background: Applied Mathematics, software engineering, digital signal processing applications, statistical process analysis, network analysis, Working experience in writing technical design specification, overseeing the development and testing process, and providing a mentorship role to junior developers and interns.

    • Knowledge of practical software engineering techniques, such as basic waterfall and incremental, extreme programming and fast prototyping.

    • Programming and Development tools: Gnu, SVN, MS Visual Studio, .NET, ATL, PERL, Mason, Various Unix shells, Purify, Quantify, Matlab
\\~\\
\textbf{PROFESSIONAL EXPERIENCE}, Full Time and Consulting

\textbf{Tableau Software, Seattle, WA						2017-present}

Senior Software Engineer 

    • Implemented Performance improvement of the existing query pipeline

    • Participated in the Implemented of classification network for error tracking and prediction
    • Migrated several server components from Windows to Linux

    • General feature development

    • Mentored and supported junior team member. 
\\~\\
\textbf{Microsoft Corp., Redmond, WA						2016-2017}

Senior Software Engineer 

    • Refactoring of the current code base

    • Design and implementation of the local Cache subsystem

    • Improved the current build and deployment process
\\~\\
\textbf{Rocket Gaming System, Las Vegas, NV						2013-2014}

Senior Math Software Engineer - Consulting

    • Design and Implementation of the Game algorithms in C++

    • Analyze and define the game characteristics, define and implement the  mathematical model

    • Refactoring, Performance and Usability improvements of the existing software and software modules.

    • Programming the Game simulations of the game math models in C++ and C\#.

    • Improved the current methodology and development process
\\~\\
\textbf{Bloomberg Inc, New York, NY							    2012-2013}

Senior Software Engineer

    • Lead the design and implementation of custom Excel function handling high frequency sell side trading and data flow. The application consists of C++  TD server, and  C\#/.NET Excel UI extension.

    • Designed and implemented the data flow, algorithms and Excel UI for the Bloomberg specific financial functions.

    • Worked with other teams to implement end to end, high frequency trading and data flow applications. The applications are the real time Excel extensions, passing large data sets to customer trading models, and executing customers orders in real time.

    • Performance improvement, refactoring of existing applications, 
\\~\\
\textbf{Translattice, Santa Clara. CA		                          2010 – 2011}

Senior Software Engineer

    • Implementation of the Paxos algorithm for distributed DB operation on Postgres DB engine in C.

    • Design and Implementation of distributed transactions,  distributed table sharding for the Postgres DB

    • Design and Implementation of the underlying multithreaded and multi process communication / replication functionality.

    • Performance improvements, refactoring existing code, Using C for the database engine and C++ for the tools.
\\~\\
\textbf{MySpace.com, Seattle, WA								2007 – 2010}

Senior Software Engineer/Architect							

    • Lead the design and implementation of MySpace Messaging system, which included internal messaging and email server and client.  The project integrated new messaging, email, comments, WebIm into the existing MySpace framework. Project was based on .NET platform, using C\#, ASP.NET, SQL Server .
    • Designed and developed message passing system supporting transaction like operation.

    • Developed Rest based services, using full .NET stack, from SQL Server as a back end DB to the C\#/.NET as middle tier.

    • Significantly refactored and the existing design, to provide the transition path to the new architecture 

    • Engineered and implemented the Web Service interoperability solutions between MySpace back end systems and external partners.

    • Providing the design for the transition to multi level application, highly scalable system.
\\~\\
textbf{Amazon.com. Seattle, WA								2004 – 2007}

Senior Software Engineer (III)								

    • Leading the design of the next generation high performance, distributed software deployment system. 

    • Designed and developed (as a lead in the team) distributed and multithread set of C++ applications to effectively manage, update and monitor over 30,000+ hosts.  The whole system consisted of several server side applications, each on separate server farm, set of client apps and daemons, persistence subsystem (Database) and a data streaming subsystem.  The amount of data stored was second only to customer data within Amazon.

    • Used several C++ libraries,  multithread, multiprocessing methodologies and applications Justified and implemented specific solutions, related to scaling, performance improvement, operational support of the software configuration, distribution and release management

    • Developed and programmed on the application server stack, from database (Oracle 9i and 10g) to application servers, Web servers, end host tools

    • Improved, refactored, and significantly increased functionality of the existing application by improving performance, availability and usability

    • Lead a team, provided designed review, support to junior team members
\\~\\
textbf{PACKETEER Inc., Kelowna, BC							2002 – 2004}

Software Engineer											

    • Implementation of compression algorithms for IP, TCP and HTTP based traffic in C.

    • Member of the design and team for of the next generation TCP traffic-processing server. The application was developed in C for kernel drivers and C++ for the higher layers.

    • Linux network stack programming and customization in C.

    • Implementation of algorithm for traffic classification, supporting XML and SOAP protocols in C++.

    • Troubleshooting of network and network based applications.

    • Design of performance critical system and application

...
\\~\\
\textbf{EDUCATION}

PhD Physics   	  (Close  to completion)					
University of Nevada Las Vegas, Las Vegas, NV

Master of Applied Mathematics						
University of Washington, Seattle, WA

Master of Engineering						
Simon Fraser University, Vancouver, BC, Canada

B.Sc. Hon in Electrical Engineering 					
School of Electrical Engineering, University of Sarajevo



